### 2014-03-14 text/Chapter00.tex
Edits in the middle of the intro chapter

\section{Coeffects: Towards context-aware languages}

The above examples cover a couple of different scenarios, but they share a common theme - 
they all talk about some \emph{context} in which an expression is evaluated. The context has 
essentially two aspects:

\begin{itemize}
\item \textbf{Flat context} represents additional data, resources and meta-data that are 
   available in the execution environment (regardless of where in the program you 
   access them). Examples include resources like GPS sensors or databases, battery status,
   framework version and similar.

\item \textbf{Structural context} contains additional meta-data related to variables. This can include
   provenance (source of the variable value), usage information (how often is the value
   accessed) or security information (does it contain sensitive data). 
\end{itemize}

As a proponent of statically typed functional languages I believe that a context-aware 
programming language should capture such context information in the type system and make
sure that basic errors (like the ones demonstrated in the four examples above) are ruled
out at compile time.

This is essentially the idea behind \emph{coeffects}. Let's look at an example showing the
idea in (a very simplified) practice and then I'll say a few words about the theory
(which is the main topic of this thesis).

@@

\subsection{Case study: Coeffects in action}

@@

\subsection{Coeffects: Theory of context dependence}

@@

\subsubsection{Brief introduction to type systems}

@@

\subsubsection{Checking what program does with effect systems}

@@

\subsubsection{Checking what program requires with coeffect systems}

>>>>

\section{Towards context-aware languages}

The four examples presented in the previous section cover different notions of \emph{context}.
The context can be viewed as execution environment, capabilities provided by the environment 
or input and meta-data about the input.

The different notions of context can be broadly classified into two categories -- those that
speak about the environment and those that speak about individual inputs (variables). In this
thesis, we refer to them as \emph{flat coeffects} and \emph{structural coeffects}, respectively:

\begin{itemize}
\item \textbf{Flat coeffects} represent additional data, resources and meta-data that are 
   available in the execution environment (regardless of how they are accessed in a program).
   Examples include resources such as GPS sensors and battery status (on a phone), databases 
   (on the server), or framework version.

\item \textbf{Structural coeffects} capture additional meta-data related to inputs. This can include
   provenance (source of the input value), usage information (how often is the value
   accessed and in what ways) or security information (whether it contain sensitive data or not). 
\end{itemize}
%
This thesis follows the tradition of statically typed programming languages. As such, we
attempt to capture such contextual information in the type system of context-aware programming
languages. The type system should provide both safety guarantees (as in the first three examples)
and also static analysis useful for optimization (as in the last example).

Although the main focus of this thesis is on the underlying theory of \emph{coeffects} and on
their structure, the following section briefly demonstrates the features that a practical
context-aware language, based on the theory of coeffects, can provide.

@@

\section{Context-aware languages in action}

@@

\section{Coeffects: Theory of context dependence}

@@

\subsection{Brief introduction to type systems}

@@

\subsection{Checking what program does with effect systems}

@@

\subsection{Checking what program requires with coeffect systems}
